{\bf {\Large Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}}

### 2015-06-04 swmp_noise.tex
incorporated Jim's comments on revision, need to proof

This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series by reducing the effect of tidal advection on continuous \ac{DO} observations.  We used a weighted regression approach originally developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.  The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.  The model is based on the recognition that daily fluctuations in \ac{DO} caused by metabolism are associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with cyclical water level changes that generally exhibit pregression relative to the solar cycle.  The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.  The method targets the tidal component as an explicit variable using dynamic model parameters that change with location in the time series.  This allows an isolation of the biological component of the \ac{DO} time series that is independent of advection.  As a result, the weighted regression approach can preserve the true biological signal rather than risking removal of both the biological and physical components as with traditional deconvulution methods.  This offers an advantage for estuaries where the magnitude of tidal effects on water quality observations can be larger than in offshore or coastal environments.  During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate the ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.  Subsequently, the simulation results informed further development of the method as applied to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).  In all examples, tidal height is used as a proxy for advection because it was recorded with the \ac{NERRS} data.  In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), the model determines the advective effect empirically.

@@

Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under a range of conditions. Specifically, the simulated data varied in the relative amount of noise in the measurement ($\varepsilon_{obs}$, $\varepsilon_{pro}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2,187 weighted regressions.

The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}. Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.  Overall, the weighted regressions produced filtered time series that were reasonable representations of the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).  The median correlation between the filtered and biological values for all time series and window widths was 0.63, with values ranging from 0.05 (very poor) to 1.00 (perfect).  Mean error was 1.02, with values ranging from 0 (perfect) to 2.12 (very poor).  Simulations with very poor performance were those those with the \ac{DO} signal composed entirely of noise from observation uncertainty. As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).  Model performance was not significantly affected by increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).  Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{tab:dtd_perf2,fig:err_surf2}).  Graphical summaries of model performance averaged by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.

@@

The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration. Anomalous values were substantially reduced for all case studies, particularly Rookery Bay and Sapelo Island.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.  However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.  The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in annual mean values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed causing a downward bias in the estimates.

@@

The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.  The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a significant amount of variation in observed \ac{DO} from tidal change.  Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location. These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series. The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.  Extension to stratified systems may be possible with additional data and model validation.

@@

For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was presumably caused by physical advection.  In particular, filtering reduced the frequency of anomalous metabolism estimates, such as negative production and positive respiration \citep{Needoba12}.  It is important to recognize, however, that `normal' estimates (positive production and negative respiration) may still include significant bias due to physical advection.  For example, a large decrease in \ac{DO} at night may appear `normal' even if inflated by a tidal effect.  In some cases, apparent tidal effects could reflect real ecosystem processes. For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime because inundation of salt marshes during the night followed by draining during a falling tide during the day lead to larger respiration values.  Resolving these patterns may simply require better data.  For the weighted regression approach, sustained periods of synchrony between the solar and tidal cycles presents a problem because collinearity between the two diminishes performance.  Moreover, \ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.  For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.  These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (\cref{fig:case_ex}).  In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the long-term seasonal variation, indicating substantial effects of physical advection.  Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.  This hypothesis should be further investigated and the method proposed herein could be used to better understand temporal variation of metabolism at different scales.

@@

A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an `end member' paradigm.  These end members are the terminal points of the water mass that is horizontally advected past the \ac{DO} sensor.  One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.  An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.  In the latter case, variation in dissolved oxygen at the \ac{DO} sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is minimal.  At sites where biological processes affect \ac{DO} at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.  Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.  Conversely, sites with mixed tides that exhibit prolonged synchronicity with the solar cycle, as at Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.  Issues of collinearity at mixed sites add additional challenges as previously noted.  Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.  For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.  We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.

@@

\caption[Example of each component of a simulated \ac{DO} time series for testing weighted regression]{Example of each component of a simulated \ac{DO} time series for testing weighted regression.  The time series were created using \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Yellow indicates a twelve hour daylight period beginning at 630 each day.\label{fig:do_sim}}
\end{figure}

@@

\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).  Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).  See \cref{tab:dtd_perf1} for a summary of combined results for each unique parameter.\label{fig:err_surf1}}
\end{figure}

@@

\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).  See \cref{tab:dtd_perf2} for a summary of combined results for each unique window.\label{fig:err_surf2}}
\end{figure}

@@

% summary of simulation performance for detided and biological, sim parameters
%latex.default(tab, file = "", where = "h", rowlabel = "Parameter",     caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,         4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,         5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",         "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")%
\begin{table}[h]
\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\varepsilon_{pro}$, $\varepsilon_{obs}$).  Values represent averages from multiple simulations with common parameters.  For example, row one is a summary of all simulations for which the diel \ac{DO} component was zero ($n=729$).  See \cref{fig:err_surf1} for results of all parameter combinations.\label{tab:dtd_perf1}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
~~0&0.05&0.20&0.41&0.86&1.00&&0.00&0.33&1.01&1.96&2.05\tabularnewline
~~1&0.28&0.44&0.61&0.93&1.00&&0.02&0.42&1.02&1.96&2.06\tabularnewline
~~2&0.55&0.60&0.81&0.96&1.00&&0.05&0.52&1.06&1.99&2.12\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
~~0&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
~~1&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
~~2&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
\hline
{\bfseries $\boldsymbol{\varepsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
~~0&0.05&0.31&0.58&0.99&1.00&&0.00&0.23&0.99&1.95&2.11\tabularnewline
~~1&0.16&0.37&0.61&0.93&0.99&&0.10&0.33&1.02&1.97&2.11\tabularnewline
~~2&0.34&0.58&0.69&0.89&0.98&&0.19&0.56&1.10&2.01&2.12\tabularnewline
\hline
{\bfseries $\boldsymbol{\varepsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
~~0&0.80&0.93&0.97&0.99&1.00&&0.00&0.17&0.29&0.46&0.84\tabularnewline
~~1&0.05&0.56&0.62&0.81&0.84&&0.97&1.00&1.02&1.08&1.28\tabularnewline
~~2&0.05&0.31&0.38&0.57&0.63&&1.93&1.97&1.98&2.01&2.12\tabularnewline
\hline
\end{tabular}\end{center}

\end{table}

% summary of simulation performance for detided and biological, window widths
%latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,     caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),     cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,     colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",         "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")%
\begin{table}[!tbp]
\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).  Values represent averages from multiple simulations with common window values. For example, row one is a summary of all simulations for which the half window width was one day ($n=729$). See \cref{fig:err_surf2} for results of all window combinations.\label{tab:dtd_perf2}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Window}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries Days}&&&&&&&&&&&\tabularnewline
~~1&0.07&0.44&0.66&0.96&1.00&&0.00&0.43&1.02&1.96&2.12\tabularnewline
~~3&0.07&0.41&0.62&0.93&1.00&&0.00&0.45&1.02&1.97&2.08\tabularnewline
~~6&0.05&0.37&0.61&0.88&1.00&&0.00&0.51&1.02&1.97&2.08\tabularnewline
\hline
{\bfseries Hours}&&&&&&&&&&&\tabularnewline
~~1&0.07&0.38&0.61&0.89&1.00&&0.00&0.51&1.02&1.97&2.05\tabularnewline
~~3&0.06&0.42&0.63&0.95&1.00&&0.00&0.43&1.02&1.97&2.05\tabularnewline
~~6&0.05&0.44&0.64&0.95&1.00&&0.00&0.48&1.04&1.96&2.12\tabularnewline
\hline
{\bfseries Tide}&&&&&&&&&&&\tabularnewline
~~0.25&0.05&0.42&0.62&0.92&1.00&&0.00&0.47&1.03&1.97&2.12\tabularnewline
~~0.5&0.07&0.42&0.63&0.94&1.00&&0.00&0.45&1.02&1.96&2.04\tabularnewline
~~1&0.07&0.42&0.63&0.94&1.00&&0.00&0.44&1.02&1.97&2.04\tabularnewline
\hline
\end{tabular}\end{center}

\end{table}

>>>>

This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series by reducing the effect of tidal advection on continuous \ac{DO} observations.  We used a weighted regression approach originally developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.  The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.  The model is based on the recognition that daily fluctuations in \ac{DO} caused by metabolism are associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with cyclical water level changes that generally exhibit pregression relative to the solar cycle.  The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.  The method targets the tidal component as an explicit variable using dynamic model parameters that change through time.  This approach makes it possible to estimate and reduce the effect of advection on the DO time series, even if this effect varies through time due to changes in metabolic rates.  Local fitting provides an advantage relative to traditional deconvulution methods. The method is most useful in estuaries where tidal effects interact with horizontal differences in metabolism and associated \ac{DO} concentrations.  Although the method could be applied in open ocean environments, it may be less useful because horizontal \ac{DO} gradients may be less in such systems, such that it could be sufficient to simply neglect advection as has commonly been done in the past. During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate the ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.  Subsequently, the simulation results informed further development of the method as applied to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).  In all examples, tidal height is used as a proxy for advection because it was recorded with the \ac{NERRS} data.  In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), the model determines the advective effect empirically.

@@

Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under different conditions. Specifically, the simulated data varied in the relative amount of noise in the measurement ($\varepsilon_{obs}$, $\varepsilon_{pro}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2,187 weighted regressions.

The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to quantify and remove the variability due to advection ($DO_{adv}$), leaving the biological response ($DO_{bio}$) (\cref{do_obs,fig:err_surf1}).  Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.  The ability of weighted regression to isolate $DO_{bio}$, indicated by high correlation and low \ac{RMSE}, was not reduced by the introduction of advection into the simulated time series (i.e., $DO_{obs}$) because weighted regression was able to quantify and remove this variability.  On the other hand, introduction of observation uncertainty ($\varepsilon_{obs}$) decreased the ability of the model to quantify and remove advection.  Perhaps counterintuitively, increased magnitude of advection improved the model performance because a larger advection signal can be fitted empirically despite the presence of random error.  When the advection signal was small, it became difficult or impossible to quantify unless there was neither process uncertainty nor observation error.  This suggests that if the advection signal is in fact very small, it would be preferable not to quantify it and instead neglect it, as has been done previously. Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{fig:err_surf2}).

@@

The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration. Anomalous values were substantially reduced for all case studies and were virtually eliminated in two of the four cases.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.  However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.  The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in annual mean values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed causing a downward bias in the estimates.

@@

The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.  The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a significant amount of variation in observed \ac{DO} from tidal change.  In application to \ac{DO} time series from \ac{NERRS} sites, the method reduces apparent interday variation in metabolic rates.  This may indicate that such interday variations result substantially from artifacts of advection, rather than real drivers of metabolic processes. This hypothesis could be investigated further by applying our proposed method while collecting measurements of metabolism and causal factors at a relatively high frequency.  The regression method and open-water technique may have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within the \ac{NERRS}. Extension to stratified systems may be possible with additional data and model development.

@@

For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was presumably caused by physical advection.  In particular, filtering reduced the frequency of anomalous metabolism estimates, such as negative production and positive respiration \citep{Needoba12}.  It is important to recognize, however, that `normal' estimates (positive production and negative respiration) may still include significant bias due to physical advection.  For example, a large decrease in \ac{DO} at night may appear `normal' even if inflated by a tidal effect.  In some cases, apparent tidal effects could reflect real ecosystem processes. For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime because inundation of salt marshes during the night followed by draining during a falling tide during the day lead to larger respiration values.  Resolving these patterns may simply require better data.  For the weighted regression approach, sustained periods of synchrony between the solar and tidal cycles presents a problem because collinearity between the two diminishes performance.  Moreover, \ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.  For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}. Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible. The open water metabolism method has generally been applied at the time scale of one day, rather than hourly. We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.

@@

Tidal currents in estuaries can move water across a mosaic of habitat types that likely have different biological rates.  For example, the magnitudes of production and respiration, and the net metabolism may differ across salinity gradients, between open water and shallow water habitats, and inside and outside of tidal marshes and seagrass habitats.  These complex estuarine seascapes are especially prevalent at sites like those in the \ac{NERRS} network.  The interaction between spatial differences in metabolic processes and tidal advection moving water masses among habitats contributes to tidal artifacts in open water metabolism calculations.  Advection alone, such as in an open coastal environment might actually generate little distortion of the metabolic signal within the \ac{DO} record.

The objective of the weighted regression method is the remove or reduce variations associated with tides, allowing for an improved estimate of the average metabolic rates between the ``end members'' that define the tidal excursion about the \ac{DO} sensor.  The method works because in most tidal regimes, tidal cycles pregress fully relative to the solar cycle within several weeks.  Pregression occurs quickly with semidiurnal tides and slightly more slowly with diurnal tides.  At time scales longer than this period, \ac{DO} variations associated with metabolism can be separated statistically from \ac{DO} variations associated with tides.  Locally-weighted regression is necessary because both the average metabolic rates associated with the solar cycle and the horizontal \ac{DO} gradients created by spatial differences in metabolism are variable in time.  The temporal scale and the choice of window width can be based on the time required for pregression of tides relative to the solar cycle.

Problems may arise in applying the method in certain situations, apparently most common with mixed tides, where complete phase pregression occurs across a seasonal time scales.  For example, \citet{Nidzieko14} showed that the higher high tide, which floods high marsh habitats, can occur at night in Elkhorn Slough throughout summer causing both real and apparent changes in metabolism.  Large errors in standard open water metabolism calculations result because tide is generally rising during afternoon and falling in the early hours of the day, amplifying the true metabolic signal.  Because the tide and solar cycle are correlated, however, our method ascribes all variability to the tide, resulting in equally unsatisfactory and large underestimates of rates.  Weighted regression could be used, but only at a very long time scale, where the benefits of using it are not as great.  As \citet{Nidzieko14} showed, these complex patterns are best quantified and understood with the benefit of a more detailed data set.

@@

\caption[Example of each component of a simulated \ac{DO} time series for testing weighted regression]{Example of each component of a simulated \ac{DO} time series for testing weighted regression.  The time series were created using \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,do_advp}. Yellow indicates a twelve hour daylight period beginning at 630 each day.\label{fig:do_sim}}
\end{figure}

@@

\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).  Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).\label{fig:err_surf1}}
\end{figure}

@@

\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).\label{fig:err_surf2}}
\end{figure}

@@

% descriptive table of case studies
%latex.default(tab, where = "!h", file = "", rowlabel = "Site",     insert.bottom = foot.val, caption = cap.val, caption.loc = "top",     cgroup = c("Tidal amplitude", "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),     n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",         "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",         "NEM"), label = "tab:case_att")%
\begin{table}[!h]
\caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$ based on observed data) for each case study.  Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).  Water quality data are averages for the entire period of record for each site.  Metabolism estimates are means of daily integrated values.\label{tab:case_att}} 
\begin{center}
\begin{tabular}{lllllcllllclll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{4}{c}{\bfseries Tidal amplitude}&\multicolumn{1}{c}{\bfseries }&\multicolumn{4}{c}{\bfseries Water quality}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Metabolism\textsuperscript{\textit{a}}}\tabularnewline
\cline{2-5} \cline{7-10} \cline{12-14}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{O1}&\multicolumn{1}{c}{P1}&\multicolumn{1}{c}{M2}&\multicolumn{1}{c}{S2}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Chl}&\multicolumn{1}{c}{Sal}&\multicolumn{1}{c}{Temp}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Pg}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
ELKVM&0.24&0.12&0.48&0.13&&7.87&3.87&32.43&13.78&&8.14&-8.19&-0.05\tabularnewline
PDBBY&0.46&0.23&0.63&0.15&&8.97&2.24&29.17&10.44&&5.95&-5.90& 0.05\tabularnewline
RKBMB&0.13&0.04&0.36&0.10&&4.48&4.50&30.53&25.85&&3.02&-3.62&-0.60\tabularnewline
SAPDC&0.10&0.02&0.54&0.07&&4.96&5.98&27.30&21.77&&4.89&-6.04&-1.16\tabularnewline
\hline
\end{tabular}\end{center}

\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism, estimated using methods described herein without removing advection effects.\end{table}

\clearpage
